Constraining gravitational wave amplitude birefringence with GWTC-3

### 2023-10-17 src/tex/ms.tex
fix bug
Update ms.tex

\begin{abstract}
    The propagation of gravitational waves can reveal fundamental features of the structure of spacetime.
    For instance, differences in the propagation of gravitational-wave polarizations would be a smoking gun for parity violations in the gravitational sector, as expected from birefringent theories like Chern-Simons gravity.
    Here we look for evidence of amplitude birefringence in the \replaced{third gravitational-wave transient}{latest Laser Interferometer Gravitational Wave Observatory-Virgo} catalog (GWTC-3) through the use of birefringent templates inspired by dynamical Chern-Simons gravity.
    From 71 binary-black-hole signals, we obtain the most precise constraints on gravitational-wave amplitude birefringence yet, measuring a birefringent attenuation of \variable{output/restricted_kappa_median.txt} at $100 \, \mathrm{Hz}$ with 90\% credibility, equivalent to a parity-violation energy scale of \variable{output/M_PV_constraint.txt}.
\end{abstract}

@@

% Section guide
In Sec.~\ref{sec:Background}, we briefly review the background of \ac{GW} amplitude birefringence.
In Sec.~\ref{sec:Method}, we describe the modification we apply to the baseline \ac{GR} waveform, summarize our \ac{PE} configuration and outline our statistical methods for combining information from multiple events.
In Sec.~\ref{sec:Results}, we present our constraint on \ac{GW} amplitude birefringence, discussing individual events and the catalog collectively.
In Sec.~\ref{sec:Discussion}, we discuss the implications of our result, comparing to previous constraints in the literature and outlining correlation structures that appear in our measurements.
Finally, we conclude in Sec.~\replaced{\ref{sec:Conclusion}}{\ref{sec:Discussion}} with a summary.

@@

To constrain birefringence, we reanalyze events from GWTC-3 \citep{GWTC-2.1, GWTC-3} implementing Eq.~\eqref{eq:waveform_modification} to directly obtain a posterior on $\kappa$ from the strain of each event.
We analyze the 71 \acp{BBH} that were detected with $\mathrm{FAR} < 1/\mathrm{yr}$, less stringent than the typical \ac{LVK} threshold of $1/1000\,\mathrm{yr}$ \cite{LIGOScientific:2020tif,LIGOScientific:2021sio}.
The \ac{FAR} values are typically determined by \ac{GR} pipelines, which could down-rank signals beyond \ac{GR} \cite{LIGOScientific:2020tif,Chia:2020psj,tgrsel}; however, since detectability through matched-filtering is most sensitive to the phase, not the amplitude, we expect only a minor decrease in sensitivity to the kind of birefringent signals explored here.
To avoid extended computations on longer signals and considering these are generally at closer distances, we do not analyze systems involving neutron stars in this work.
We procure strain data from \add{the} Gravitational Wave Open Science Center \citep{GWOSC}.

@@

The result of \citet{Okounkova_2022} is also phrased in terms of a canonical Chern-Simons length scale,\footnote{This results from setting the so-called ``canonical'' Chern-Simons embedding into an Friedmann-Robertson-Walker cosmology, where the Chern-Simons field evolves as $\theta \propto t$ \cite{Alexander:2009tp,Jackiw:2003pm,Yunes2009}.}
\begin{equation}
    l_0 = \frac{c d_H \kappa}{3 \pi f} = 1400 \, \mathrm{km} \left( \frac{\kappa}{1 \, \mathrm{Gpc}^{-1}} \right) \left( \frac{100 \, \mathrm{Hz}}{f} \right);
\end{equation}
our 68\% credible bound on $\kappa$ implies $l_0 \lesssim 40 \, \mathrm{km}$ at $100 \, \mathrm{Hz}$.  This compares favorably to the constraint in \citet{Okounkova_2022}, which had $l_0 \lesssim 1000 \, \mathrm{km}$.

@@

\section{EXTENDED RESULTS FOR NOTABLE EVENTS}

>>>>

\begin{abstract}
    The propagation of gravitational waves can reveal fundamental features of the structure of spacetime.
    For instance, differences in the propagation of gravitational-wave polarizations would be a smoking gun for parity violations in the gravitational sector, as expected from birefringent theories like Chern-Simons gravity.
    Here we look for evidence of amplitude birefringence in the \replaced{third catalog of detections by the Laser Interferometer Gravitational Wave Observatory and Virgo}{latest Laser Interferometer Gravitational Wave Observatory-Virgo catalog (GWTC-3)} through the use of birefringent templates inspired by dynamical Chern-Simons gravity.
    From 71 binary-black-hole signals, we obtain the most precise constraints on gravitational-wave amplitude birefringence yet, measuring a birefringent attenuation of \variable{output/restricted_kappa_median.txt} at $100 \, \mathrm{Hz}$ with 90\% credibility, equivalent to a parity-violation energy scale of \variable{output/M_PV_constraint.txt}.
\end{abstract}

@@

% Section guide
In Sec.~\ref{sec:Background}, we briefly review the background of \ac{GW} amplitude birefringence.
In Sec.~\ref{sec:Method}, we describe the modification we apply to the baseline \ac{GR} waveform, summarize our \ac{PE} configuration and outline our statistical methods for combining information from multiple events.
In Sec.~\ref{sec:Results}, we present our constraint on \ac{GW} amplitude birefringence, discussing individual events and the catalog collectively.
In Sec.~\ref{sec:Discussion}, we discuss the implications of our result, comparing to previous constraints in the literature and outlining correlation structures that appear in our measurements.
Finally, we conclude in Sec.~\replaced{\ref{sec:Conclusion}}{\ref{sec:Discussion}} with a summary.
\added{Additional details on the results for some notable events are presented in \ref{sec:Appendix}.}

@@

To constrain birefringence, we reanalyze events from GWTC-3 \citep{GWTC-2.1, GWTC-3} implementing Eq.~\eqref{eq:waveform_modification} to directly obtain a posterior on $\kappa$ from the strain of each event.
We analyze the 71 \acp{BBH} that were detected with $\mathrm{FAR} < 1/\mathrm{yr}$, less stringent than the typical \ac{LVK} threshold of $1/1000\,\mathrm{yr}$ \cite{LIGOScientific:2020tif,LIGOScientific:2021sio}.
The \ac{FAR} values are typically determined by \ac{GR} pipelines, which could down-rank signals beyond \ac{GR} \cite{LIGOScientific:2020tif,Chia:2020psj,tgrsel}; however, since detectability through matched-filtering is most sensitive to the phase, not the amplitude, we expect only a minor decrease in sensitivity to the kind of birefringent signals explored here.
To avoid extended computations on longer signals and considering these are generally at closer distances, we do not analyze systems involving neutron stars in this work.
We procure strain data from Gravitational Wave Open Science Center \citep{GWOSC}.

@@

The result of \citet{Okounkova_2022} is also phrased in terms of a canonical Chern-Simons length scale,\footnote{This results from setting the so-called ``canonical'' Chern-Simons embedding into a Friedmann-Robertson-Walker cosmology, where the Chern-Simons field evolves as $\theta \propto t$ \cite{Alexander:2009tp,Jackiw:2003pm,Yunes2009}.}
\begin{equation}
    l_0 = \frac{c d_H \kappa}{3 \pi f} = 1400 \, \mathrm{km} \left( \frac{\kappa}{1 \, \mathrm{Gpc}^{-1}} \right) \left( \frac{100 \, \mathrm{Hz}}{f} \right);
\end{equation}
our 68\% credible bound on $\kappa$ implies $l_0 \lesssim 40 \, \mathrm{km}$ at $100 \, \mathrm{Hz}$.  This compares favorably to the constraint in \citet{Okounkova_2022}, which had $l_0 \lesssim 1000 \, \mathrm{km}$.

@@

\section{EXTENDED RESULTS FOR NOTABLE EVENTS}
\label{sec:Appendix}
